---
title: "Predictive Modeling Process"
author: "Shirley Jin, Xiaoqian Zhu"
date: "November 4, 2016"
output: ioslides_presentation
---
```{r, echo = FALSE, include = FALSE, eval=TRUE}
# load the necessary data files and required packages
scaled_credit <- read.csv("../data/datasets/scaled-credit.csv")
load("../data/output/ols-regression.RData")
load("../data/output/lasso-regression.RData")
load("../data/output/ridge-regression.RData")
load("../data/output/pc-regression.RData")
load("../data/output/pls-regression.RData")

# packages
library(xtable)
library(png)
library(grid)
library(ggplot2)
library(reshape)
library(Matrix)
```

## Introduction
- In this project, we will recreate the predictive modeling process in Chapter 6: Linear Model Selection and Regularization of the book [**An Introduction to Statistical Learning**](http://www-bcf.usc.edu/~gareth/ISL/).
- We will be using multiple regression models on the _Credit_ dataset and compare the differences between these models.


## Data

- Variables
    - Predictor Variables
        - quantitative: age, cards, income, limit, education, and rating
        - qualitative: gender, student, status, ethnicity
    - Response Variable
        - quantitative: balance
        

## Pre-Modeling Data Processing
There are two major processing steps we took to prepare the data:

- Dummy out categorical variables
    - Transforming categorical variable into dummy variables
```
temp_credit <- model.matrix(Balance ~ ., data = credit)
new_credit <- cbind(temp_credit[ ,-1], Balance = credit$Balance)
```


- Mean centering and standardizing
    - each variable will have mean zero, and standard deviation one to have comparable scales
```
scaled_credit <- scale(new_credit, center = TRUE, scale = TRUE)
```


## Preparing Training and Testing Data
In order to evaluate the models, we will take a random sample of the data for model building and use the rest to understand the model performance.

- **training set**: generate a random sample of size 300. Use sample() function and an integer vector 1:400 to sample 300 values
- **testing set**: rest of 100 observations not included in the train set
  
  
## Regression Models
- Ordinary Least Squares Regression (OLS)
- Shrinkage Methods
    - Ridge Regression
    - Lasso Regression
- Dimension Reduction Methods
    - Principal Components Regression (PCR)
    - Partial Least Squares Regression (PLSR)
  
  
## OLS

- Minimizing the RSS in order to minimize the error of the prediction

$$Sales \approx \beta_0 + \beta_1*Income+ \beta_2*Limit + \dots + \beta_11*EthnicityCaucasian$$


## Ridge Regression

- Shrinkage method
- Add a tuing parameter based on OLS
- Flexibility of ridge regression decreases with increases in tuning parameter

$$RSS + \lambda\sum_{j=1}^{p}\beta_j^{2}$$


## Lasso Regression

- Shinkage method
- Will result in estimations of coefficients to zero

$$\sum_{i=1}^{n}(y_i-\beta_0-\sum_{j=1}^{p}\beta_jx_{ij})^2+\lambda\sum_{j=1}^{p}|\beta_j|$$


## PCR

PCR is a dimension reduction technique for regression that involves constructing the first $M$ principal components, $Z_1 + \dots + Z_{M*}$, and then using these components as the predictors in a linear regression model that is fit using least squares.


## PLSR

- An extension of PCR

$$Z_1=\sum_{j=1}^p\phi_{j1}X_j$$

## Analysis
## OLS

This is coefficient table for OLS:
```{r results= 'asis', echo = FALSE, error=FALSE}
library(Matrix)
library(xtable)
load("../data/output/ols-regression.RData")
print(xtable(ols_summary$coefficients, caption = 'OLS Coefficients',digits = c(0,5,5,5,5)), comment = FALSE,type = 'latex')
```

## OLS

- Benchmark for comparison
- Education, Gender, Martial Status and Ethnicity has big p-value， thus not statitiscally significant.
- Income, Limit, and Rating have large and statistically significant coefficients.

## Ridge

This is the coefficient table for Ridge:
```{r results= 'asis', echo = FALSE, error=FALSE}
load("../data/output/ridge-regression.RData")
ridge_matrix_coef <- as.matrix(ridge_coefficients)
colnames(ridge_matrix_coef) <- "Estimates"
print(xtable(ridge_matrix_coef, caption = 'Ridge Coefficients',digits = c(0,5)), comment = FALSE,type = 'latex')
```



## Ridge

- The smallest validation error is $\lambda$ = `r ridge_min_lambda`.
- Small $\lambda$
- Similar coefficients with OLS, expect the larger Rating coefficients.


## Ridge

This is the image for ridge cross validation errors.
$$\includegraphics[width=250pt]{../images/ridge-cross-validation-errors.png}$$


## Lasso

This the coefficient table for Ridge:
```{r results= 'asis', echo = FALSE, error=FALSE}
load("../data/output/lasso-regression.RData")
lasso_matrix_coef <- as.matrix(lasso_coefficients)
colnames(lasso_matrix_coef) <- "Estimates"
print(xtable(lasso_matrix_coef, caption = 'Lasso Coefficients', digits = c(0,5)), comment = FALSE,type = 'latex')
```


## Lasso

- $\lambda$ = `r lasso_min_lambda`
- Six of the coefficients are set to zero
- Smaller coefficients than OLS

## Lasso

The graph for lasso cross-validation errors:
$$\includegraphics[width=250pt]{../images/lasso-cross-validation-errors.png}$$


## PCR

This is the coefficient table for PCR:
```{r results= 'asis', echo = FALSE, error=FALSE}
load("../data/output/pc-regression.RData")
pcr_matrix_coef <- as.matrix(pc_coef_full)
dimnames(pcr_matrix_coef) <- list(rownames(pcr_matrix_coef, do.NULL = FALSE, prefix = "row"),colnames(pcr_matrix_coef, do.NULL = FALSE, prefix = "col"))
colnames(pcr_matrix_coef) <- "Estimates"
rownames(pcr_matrix_coef) <- c("Income", "Limit", "Rating", "Cards", "Age", "Education", "GenderFemale", "StudentYes", "MarriedYes", "EthnicityAsian", "EthnicityCaucasian")
print(xtable(pcr_matrix_coef, caption = 'PCR Coefficients',digits = c(0,5)), comment = FALSE,type = 'latex')
```


## PCR

- The best M is `r lambda_min_pc`.
- Almost same coefficients with OLS and PLSR.
- Coefficients also similar to Ridge and Lasso.

## PCR

This is the MSE plot for PCR:
$$\includegraphics[width=250pt]{../images/cv-pc-mse-plot.png}$$

## PLSR

This is coefficient table for PLSR:
```{r results= 'asis', echo = FALSE, error=FALSE}
load("../data/output/pls-regression.RData")
pls_matrix_coef <- as.matrix(pls_coef_full)
dimnames(pls_matrix_coef) <- list(rownames(pls_matrix_coef, do.NULL = FALSE, prefix = "row"),colnames(pls_matrix_coef, do.NULL = FALSE, prefix = "col"))
colnames(pls_matrix_coef) <- "Estimates"
rownames(pls_matrix_coef) <- c("Income", "Limit", "Rating", "Cards", "Age", "Education", "GenderFemale", "StudentYes", "MarriedYes", "EthnicityAsian", "EthnicityCaucasian")
print(xtable(pls_matrix_coef, caption = 'PLS Coefficients',digits = c(0,5)), comment = FALSE,type = 'latex')
```


## PLSR

- The best M is `r lambda_min_pls`.
- Almost same to OLS and PCR.
- Similar to Ridge.


## PLSR

This is the MSE plot for PLSR:
$$\includegraphics[width=250pt]{../images/cv-pls-mse-plot.png}$$


## Results

This is the table for MSE of all regression models:
```{r results= 'asis', echo = FALSE}
MSE_five = as.matrix(c("ols" = ols_mse, "ridge" = ridge_mse, "lasso" = lasso_mse, "pcr" = pc_mse, "plsr" = pls_mse))
colnames(MSE_five) = "MSE"
print(xtable(MSE_five, caption = 'MSE of Five Regression Methods', digits = 5), comment = FALSE)
```


## MSE

- Close to one another
- lasso has the largest MSE `r lasso_mse`, ols the least `r ols_mse`
- ols the best fit model, lasso the worst fit model

## Results

This the trendline plot for each type of regression models fit on Credit.csv:
```{r results= 'asis', echo = FALSE}
Variables <- c("Intercept", "Income","Limit", "Rating", "Cards", "Age", "Education", "GenderFemale", "StudentYes", "MarriedYes", "EthnicityAsian", "EthnicityCaucasian")
OLS <- as.vector(ols_coefficients)
Ridge <- as.vector(ridge_coefficients)
Lasso <- as.vector(lasso_coefficients)
PCR <- as.vector(pc_coef_full)
PCR <- append(PCR, 0, 0)
PLSR <- as.vector(pls_coef_full)
PLSR <- append(PLSR, 0, 0)

five_matrix_coef <- data.frame(Variables, OLS, Ridge, Lasso, PCR, PLSR)
print(xtable(five_matrix_coef, caption = 'Regression Coefficients for 5 Regression Methods', digits = 3), comment = FALSE, type = 'latex')
```


## Official Coefficients

- Lasso with largest variation
- PLSR, OLSR and PCR have the almost same coefficients


## Conclusions

- OLS is the best fit model to Credit dataset.
- Income, Limit, and Rating are statistically and economiclly significant across five models.
- Lasso is the worst fit model to Credit dataset.
